The Bot with the Eloquent Speech: Exploring the Relationship between Perceptions of Conversational Agents and Their Conversation Architecture

### 2023-02-14 paper.tex
title & abstract updates, update figures, added coverage heatmap

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
% \title[short title]{How to Make Pinocchio a Real Boy? Designing conversational architecture elements to achieve desirable anthropomorphized perceptions}
\title[Bot's Guide to Being Human]{Bot's Guide to Being Human: Investigating the Effect of Conversation Architecture Elements on the Anthropomorphized Perceptions of Conversational Agents}
% Blurring the line between human and bots

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
The natural language interactions of conversational agents encourage users to attribute humanlike qualities to them, but we don't yet have a systematic understanding on how to design conversation architecture to orchestrate these anthropomorphized perceptions. Through a comprehensive review on papers with empirical studies on the impact of conversation architecture elements on perception of agents, we collected explored connections in literature to create a consistent taxonomy for anthropomorphized perceptions, and conversation architecture elements. Also, we created a generalized framework of the relationship between conversation architecture and perceptions based on the number of impacts found in the reviewed studies. In our analysis, we found there are inconsistencies in the measurements for perceptions of agents, as well as influencing factors other than conversation architecture that could impact anthropomorphized perceptions of agents. Also, it is important to consider ethical aspects of designing anthropomorphized perceptions of CAs as it may have negative consequences to users.
\end{abstract}

@@

@@

\textbf{Category} & \textbf{Theme} & \textbf{Sample Codes} & \textbf{No. Papers} & \textbf{Papers} 

@@

\begin{table*}[h]
\resizebox{\textwidth}{!}{
\renewcommand*{\arraystretch}{1.4}
\begin{tabular}{@{}p{0.17\textwidth} | p{0.14\textwidth} | p{0.25\textwidth} | >{\centering}p{0.09\textwidth} | p{0.34\textwidth} @{}}
\Xhline{1.2pt}
\textbf{Category} & \textbf{Theme} & \textbf{Sample Codes} & \textbf{No. Papers} & \textbf{Papers}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Content}    & Affective Language & emotional expressions, sentiment-adaptive responses & 11
& \cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{lee2019s}\cmt{[55]}\cite{lubis2019positive}\cmt{[43]}\cite{moilanen2022measuring}\cmt{[82]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Social Talk & self-disclosure, social dialogue & 7
& \cite{andrews2012system}\cmt{[38]}\cite{lee2020hear}\cmt{[23]}\cite{lubold2016effects}\cmt{[86]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Humour & humorous content, jokes & 5
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Agent-Initiated Content & proactive dialogue, conversation repair & 4
& \cite{ashktorab2019resilient}\cmt{[88]}\cite{cuadra2021my}\cmt{[67]}\cite{kraus2020effects}\cmt{[64]}\cite{xiao2021let}\cmt{[73]}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Style} & Formality & formal vs. casual, honorific expressions & 9
& \cite{cox2022does}\cmt{[27]}\cite{elsholz2019exploring}\cmt{[61]}\cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2019comparing}\cmt{[89]}\cite{ma2022ask}\cmt{[29]}\cite{moilanen2022measuring}\cmt{[82]}\cite{ouchi2019should}\cmt{[59]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Alignment & content matching, lexical alignment, agreeableness & 7
& \cite{healey2013relating}\cmt{[39]}\cite{hoegen2019end}\cmt{[31]}\cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}\cite{spillner2021talk}\cmt{[18]}\cite{volkel2021examining}\cmt{[69]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Elaborateness & elaborate, concise, sentence structure & 6
& \cite{haas2022keep}\cmt{[78]}\cite{miehle2018exploring}\cmt{[51]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Disfluency & fillers, interjections, repetitions & 6
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{wester2015artificial}\cmt{[14]}\cite{yang2021effect}\cmt{[72]}
\\ \Xhline{1.2pt}
\multirow{2}{*}{Non-Linguistic Format} & Prosodic Cue & pitch, intonation, speech rate, spoken accent & 12
& \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{feijoo2021effects}\cmt{[70]}\cite{habler2019effects}\cmt{[63]}\cite{hu2021enhancing}\cmt{[56]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2020can}\cmt{[24]}\cite{lubold2016effects}\cmt{[86]}\cite{misu2011toward}\cmt{[83]}\cite{tolmeijer2021female}\cmt{[62]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Text Variation & typo, capitalization, emoticons, response delay & 9
& \cite{fadhil2018effect}\cmt{[52]}\cite{gnewuch2018faster}\cmt{[19]}\cite{gnewuch2018chatbot}\cmt{[21]}\cite{gnewuch2022opposing}\cmt{[20]}\cite{kim2019comparing}\cmt{[89]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{westerman2019believe}\cmt{[9]}\cite{wilhelm2022keep}\cmt{[28]}
\\ \Xhline{1.2pt}
\end{tabular}%
}
\caption{Conversation architecture elements}
\label{tab:cues}
\end{table*}

\subsection{Relationship Between Anthropomorphized Perceptions and Conversation Architecture Elements}

@@

>>>>

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
% \title[short title]{How to Make Pinocchio a Real Boy? Designing conversational architecture elements to achieve desirable anthropomorphized perceptions}
\title[The Bot with the Eloquent Speech]{The Bot with the Eloquent Speech: Exploring the Relationship between Perceptions of Conversational Agents and Their Conversation Architecture}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
Previous research has demonstrated a relationship between the structure of conversational agents' natural language interaction and how these agents are perceived by their users. However, the systematic connections between the perception of agents and elements of conversation architecture are not well understood. Furthermore, there is no clear definition of the aspects of perception that are affected and the elements of conversation architecture. In this paper, we present a synthesized framework demonstrating the connections between perceptions of agents and conversation architecture elements, using a taxonomy of eleven perceptions grouped into four categories, and ten elements grouped into three categories. We highlight the need for further research for the perception of agent's ability and perception of social connection with agent's, especially as it relates to the linguistic style of conversational agents.  Also, there is the need to be consistent in the assessment of agents' perceptions, and address ethical concerns in design.
\end{abstract}

@@

%%%% TODO %%%%
% Add headmap of explored relationships - side by side with the connections - show numbers - DONE
% Put the heatmap of impact earlier in the paper - DONE
% Put the tables together - DONE
% show what is explored vs. the connections that had impacts. - DONE
% User perception of agent instead of anthropomorphized perceptions
% Refer to the explored connection in writing, and connections with impacts
% Update to use perception aspects and categories (instead of themes and organizing themes)
% Update introduction
% Include focus areas for future reasearch
% Reviewed recording up to 17:31
% Question about single column vs. double column format - ask Jaisie

@@

\multicolumn{2}{l|}{\textbf{Perception Aspects}} & \textbf{Examples} & \textbf{No. Papers} & \textbf{Papers} 

@@

\begin{table*}[h]
\resizebox{\textwidth}{!}{
\renewcommand*{\arraystretch}{1.4}
\begin{tabular}{@{}p{0.17\textwidth} | p{0.14\textwidth} | p{0.25\textwidth} | >{\centering}p{0.09\textwidth} | p{0.34\textwidth} @{}}
\Xhline{1.2pt}
\textbf{Element Category} & \textbf{Element} & \textbf{Examples} & \textbf{No. Papers} & \textbf{Papers}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Content}    & Affective Language & emotional expressions, sentiment-adaptive responses & 11
& \cite{daher2020empathic}\cmt{[58]}\cite{diederich2019emulating}\cmt{[25]}\cite{healey2013relating}\cmt{[39]}\cite{hu2022polite}\cmt{[76]}\cite{lee2019s}\cmt{[55]}\cite{lubis2019positive}\cmt{[43]}\cite{moilanen2022measuring}\cmt{[82]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{yang2017perceived}\cmt{[44]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Social Talk & self-disclosure, social dialogue & 7
& \cite{andrews2012system}\cmt{[38]}\cite{lee2020hear}\cmt{[23]}\cite{lubold2016effects}\cmt{[86]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Humour & humorous content, jokes & 5
& \cite{ceha2021can}\cmt{[57]}\cite{go2021conversational}\cmt{[80]}\cite{khooshabeh2011does}\cmt{[37]}\cite{miyamoto2017improving}\cmt{[46]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Agent-Initiated Content & proactive dialogue, conversation repair & 4
& \cite{ashktorab2019resilient}\cmt{[88]}\cite{cuadra2021my}\cmt{[67]}\cite{kraus2020effects}\cmt{[64]}\cite{xiao2021let}\cmt{[73]}
\\ \Xhline{1.2pt}
\multirow{4}{*}{Linguistic Style} & Formality & formal vs. casual, honorific expressions & 9
& \cite{cox2022does}\cmt{[27]}\cite{elsholz2019exploring}\cmt{[61]}\cite{habler2019effects}\cmt{[63]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2019comparing}\cmt{[89]}\cite{ma2022ask}\cmt{[29]}\cite{moilanen2022measuring}\cmt{[82]}\cite{ouchi2019should}\cmt{[59]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Alignment & content matching, lexical alignment, agreeableness & 7
& \cite{healey2013relating}\cmt{[39]}\cite{hoegen2019end}\cmt{[31]}\cite{huiyang2022improving}\cmt{[17]}\cite{linnemann2018can}\cmt{[15]}\cite{spillner2021talk}\cmt{[18]}\cite{volkel2021examining}\cmt{[69]}\cite{volkel2021manipulating}\cmt{[68]}
\\ \cline{2-5}
& Elaborateness & elaborate, concise, sentence structure & 6
& \cite{haas2022keep}\cmt{[78]}\cite{miehle2018exploring}\cmt{[51]}\cite{moilanen2022measuring}\cmt{[82]}\cite{roy2021users}\cmt{[71]}\cite{volkel2021manipulating}\cmt{[68]}\cite{volkel2022user}\cmt{[75]}
\\ \cline{2-5}
& Disfluency & fillers, interjections, repetitions & 6
& \cite{ceha2022expressive}\cmt{[77]}\cite{hu2021enhancing}\cmt{[56]}\cite{jeong2019exploring}\cmt{[10]}\cite{niewiadomski2013laugh}\cmt{[85]}\cite{wester2015artificial}\cmt{[14]}\cite{yang2021effect}\cmt{[72]}
\\ \Xhline{1.2pt}
\multirow{2}{*}{Non-Linguistic Format} & Prosodic Cue & pitch, intonation, speech rate, spoken accent & 12
& \cite{chan2021kinvoices}\cmt{[74]}\cite{choi2020nobody}\cmt{[54]}\cite{dubiel2020persuasive}\cmt{[60]}\cite{feijoo2021effects}\cmt{[70]}\cite{habler2019effects}\cmt{[63]}\cite{hu2021enhancing}\cmt{[56]}\cite{jestin2022effects}\cmt{[81]}\cite{kim2020can}\cmt{[24]}\cite{lubold2016effects}\cmt{[86]}\cite{misu2011toward}\cmt{[83]}\cite{tolmeijer2021female}\cmt{[62]}\cite{zhu2022effects}\cmt{[26]}
\\ \cline{2-5}
& Text Variation & typo, capitalization, emoticons, response delay & 9
& \cite{fadhil2018effect}\cmt{[52]}\cite{gnewuch2018faster}\cmt{[19]}\cite{gnewuch2018chatbot}\cmt{[21]}\cite{gnewuch2022opposing}\cmt{[20]}\cite{kim2019comparing}\cmt{[89]}\cite{seeger2021chatbots}\cmt{[35]}\cite{volkel2022user}\cmt{[75]}\cite{westerman2019believe}\cmt{[9]}\cite{wilhelm2022keep}\cmt{[28]}
\\ \Xhline{1.2pt}
\end{tabular}%
}
\caption{Conversation architecture elements}
\label{tab:cues}
\end{table*}

\subsection{Relationship Between Perceptions of Agents and Conversation Architecture}

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=\textwidth]{fig-heatmap-coverage.png}
  \caption{Heatmap of literature for perception of conversational agents for each type of conversational architecture element}
  \label{fig:heatmap-coverage}
\end{figure*}
